---
title: "IXIS Data Science Challenge"
author: "Joran Elias"
format: beamer
fontsize: 8pt
---

```{r}
#| label: setup
#| echo: false
#| messages: false
#| include: false
library(tidyverse)
library(janitor)
library(skimr)
library(lubridate)
library(writexl)
library(lme4)
library(patchwork)

theme_set(theme_bw(base_size = 18) + 
            theme(legend.direction = 'horizontal',
                  legend.position = 'bottom'))

# Read data
session_counts <- readr::read_csv(file = '../data/DataAnalyst_Ecom_data_sessionCounts.csv',
                                  show_col_types = FALSE) %>%
  janitor::clean_names()
adds_to_cart <- readr::read_csv(file = '../data/DataAnalyst_Ecom_data_addsToCart.csv',
                                show_col_types = FALSE) %>%
  janitor::clean_names()

# Ensure dim_date is a date
session_counts <- session_counts %>%
  mutate(dim_date = as.Date(dim_date,'%m/%d/%y'))

# Create a true monthly date column to ease filtering
adds_to_cart <- adds_to_cart %>%
  mutate(month = as.Date(paste(dim_year,
                               stringr::str_pad(dim_month,width = 2,side = 'left',pad = '0'),
                               '01',sep = '-'))) %>%
  relocate(month,.before = everything()) 

# Mobile & tablet is basically Safari vs everything else, desktop is more
# diverse
session_counts <- session_counts %>%
  mutate(dim_browser_cat = case_when(
    grepl('Safari',dim_browser) ~ 'Safari',
    dim_device_category %in% c('mobile','tablet') & !grepl('Safari',dim_browser) ~ 'Other',
    dim_browser %in% c('Chrome','Internet Explorer','Firefox') ~ dim_browser,
    TRUE ~ 'Other'
  ))

# Month * Device aggregation with the metrics:
# - Session
# - Transactions
# - QTY
# - ECR (Transactions / Sessions)
# 
# month_idx is for use in regression model to ensure sensible interpretation
# of coef (i.e. each unit change is one month)
month_device_metrics <- session_counts %>%
  mutate(month = lubridate::floor_date(dim_date,'month'),
         month_idx = as.integer(as.factor(month))) %>%
  group_by(month,month_idx,dim_device_category) %>%
  summarise(across(where(is.numeric),sum),.groups = 'drop') %>%
  mutate(ecr = transactions / sessions)
```

## Conversion Rates By Device Category
```{r}
#| include: false

# Simple varying slopes model to quantify trends for each category
dev_m <- lme4::lmer(formula = ecr ~ month_idx + (1+month_idx|dim_device_category),
              data = month_device_metrics,
              REML = FALSE)

dev_slopes <- coef(dev_m)[[1]]$month_idx * 100
```
:::: {.columns}

::: {.column width="25%"}

- Conversion rates are best on desktop devices, worst on mobile
- Desktop & mobile conversion rates have trended gradually up (`r format(dev_slopes[1],digits = 2)` \& `r format(dev_slopes[2],digits = 2)` per month respectively)
- Tablet conversion rate has trended gradually down (`r format(dev_slopes[3],digits = 2)` per month)

:::

::: {.column width="75%"}
```{r}
month_device_metrics %>%
  mutate(dim_device_category = factor(dim_device_category,levels = c('desktop','tablet','mobile'))) %>%
ggplot(data = .,
       aes(x = month,y = ecr,color = dim_device_category)) + 
  geom_line() + 
  geom_smooth(formula = y~x,method = 'lm',se = FALSE,linetype = 'dashed') + 
  scale_y_continuous(labels = scales::percent_format()) + 
  scale_color_brewer(type = 'qual') +
  labs(x = 'Month',y = 'ECR',color = 'Device Category')
```
:::

::::

## Conversion Rates - Desktop & Mobile
```{r}
#| include: false
month_dev_brow_ecr_p <- session_counts %>%
  mutate(month = lubridate::floor_date(dim_date,'month')) %>%
  group_by(month,dim_device_category,dim_browser_cat) %>%
  summarise(across(where(is.numeric),sum),.groups = 'drop') %>%
  mutate(ecr = transactions / sessions) %>%
  select(month,dim_device_category,dim_browser_cat,ecr)
```

:::: {.columns}
::: {.column width="25%"}

- Desktop conversion has risen overall, but the best converting browser category, Safari, has declined slightly
- Mobile conversion rates are small, but strong growth on mobile Safari

:::

::: {.column width="75%"}

```{r}
#| warnings: false
#| fig.height: 10
p_mobile <- month_dev_brow_ecr_p %>% 
  filter(dim_device_category == 'mobile') %>%
  mutate(dim_browser_cat = factor(dim_browser_cat,levels = c('Safari','Other'))) %>%
  ggplot(data = .,aes(x = month,y = ecr,color = dim_browser_cat)) +
  facet_wrap(~dim_device_category) +
  geom_line() + 
  geom_smooth(formula = y~x,method = 'lm',se = FALSE,linetype = 'dashed') +
  scale_y_continuous(labels = scales::percent_format()) + 
  scale_color_brewer(type = 'qual') +
  labs(x = 'Month',y = 'ECR',color = 'Browser Category')

p_desktop <- month_dev_brow_ecr_p %>% 
  filter(dim_device_category == 'desktop') %>%
  ggplot(data = .,aes(x = month,y = ecr,color = dim_browser_cat)) +
  facet_wrap(~dim_device_category) +
  geom_line() + 
  geom_smooth(formula = y~x,method = 'lm',se = FALSE,linetype = 'dashed') +
  scale_y_continuous(labels = scales::percent_format()) + 
  scale_color_brewer(type = 'qual') +
  labs(x = 'Month',y = 'ECR',color = 'Browser Category') + 
  guides(color = guide_legend(nrow = 2))

# Combine plots vis patchwork
p_desktop / p_mobile
```
:::
::::

## Conversion Rates - Tablet
:::: {.columns}
::: {.column width="25%"}

- Apparent small decline in tablet conversion rates due to single large value in July 2012 on mobile Safari 
- Without this potential outlier, tablet conversion would likely appear flat for the year

:::

::: {.column width="75%"}

```{r}
month_dev_brow_ecr_p %>% 
  filter(dim_device_category == 'tablet') %>%
  mutate(dim_browser_cat = factor(dim_browser_cat,levels = c('Safari','Other'))) %>%
  ggplot(data = .,aes(x = month,y = ecr,color = dim_browser_cat)) +
  geom_line() + 
  geom_smooth(formula = y~x,method = 'lm',se = FALSE,linetype = 'dashed') +
  scale_y_continuous(labels = scales::percent_format()) + 
  scale_color_brewer(type = 'qual') +
  labs(x = 'Month',y = 'ECR',color = 'Browser Category')
```
:::
::::
## Recommendations

::: {.incremental}

- Focus on mobile Safari as a high growth opportunity
- Investigate improvements to iPad Safari experience
- Investigate improvements to desktop Safari experience

:::
